---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
#Do not change this Chunk

knitr::opts_chunk$set(echo = FALSE, error = TRUE)

library(tidyverse)
library(ggplot2)
library(DBI)




#enable other connections
# dbDisconnect(con)
#
# connect to CapeCodd
congap <- DBI::dbConnect(RSQLite::SQLite(), "gapminderdb.db")
conbook <- DBI::dbConnect(RSQLite::SQLite(), "books.db")
#


```
## Before you start

Make sure that you have copied the StudentNoteBook to a folder on you computer, as well as the 2 database files. The 2 database files need to be in the same folder as the StudentNoteBook. 

You can test the locataion by running the following code

```{r eval=FALSE, echo=TRUE}
getwd()
```

Make sure it returns your Exam_HandIn folder. If it does not return, save your StudentNoteBook in the Exam_HandIn folder. If you know that you are saving in the Exam_HandIn folder you can continue. 

Now add your name:

Name: Valentin Vasilev

Studentnumber:3860140

Class:  PC05


## Introduction text

In this exam we will go over the theory you worked with earlier. In four sections we will focus on some of the questions and theory presented in the course of DDO.
We will start with some questions regarding reading the ERD. Then there are 2 datasets in which we will run several queries, the first you might recognize as the Gapminder, the second is regarding different authors. Lastly we will combine the Queries that you have made with the Tidyverse skills we have obtained earlier in the course.  

As we will be using different libraries you need to enable these:

```{r, eval=FALSE, echo=TRUE}
library(tidyverse)
library(ggplot2)
library(DBI)
```

## When finished

After completing all questions you can save the Notebook. Using the <i> "knit" </i>function to print your work into an HTML file completes the last task of the test. When you want to upload the work with EFIT make sure to close all applications, otherwise you cannot upload. 

When everything is uploaded you can leave the room quietly. 


## Understanding a database

For this assignemnt we have the following ERD: 

<img src = "images/car.png" />

in this ERD there are 6 relationships, for the relations would the following be true or false:

1. A vehicle will always have a rental

>FALSE
   
2. A Rental will always have 1 Reservation

>TRUE

3. A vehicle is listed with one office only

>TRUE

4. For every vehicle type there is at least 1 vehicle

>FALSE

5. Every table will always need a primary key

>TRUE

6. For a reservation it is optional to have a VehicleType

>FALSE

Describe the following relationships in plain English:

7. The relationship between Office and Vehicle - There is an office per vehicle

8. The relationship between Customer and Reservation - One customer can have many reservations

9. The relationship between Rental and Vehicle - Every Rental always has a vehicle

## gapminder queries.

In week 6 we have worked with Gapminder and the tidyverse. Now for the SQL we have the gapminder database available. To see the content of the dataset we can use the following code:

```{sql connection=congap}
SELECT *
FROM gapminder Where country = "Afghanistan" and year < 2000;
```



10. Show the name of countries, year, and life expectation from the table gapminder.

```{sql connection=congap}
SELECT country, year, lifeExp
FROM gapminder
```



11. List all the countries from the gapminder table. Make sure there are no duplicate lines.

```{sql connection=congap}
SELECT DISTINCT country
FROM gapminder
```



12. List the information of China in 1967 from the Gapminder table

```{sql connection=congap}
SELECT *
FROM gapminder
WHERE country = "China" AND year = 1967
```



13. What is the total population in 2007 according to the gapminder table

```{sql connection=congap}
SELECT SUM(pop) as totalPopIn2007
FROM gapminder
WHERE year = 2007
```



14. Calculate the average gdpPercap of each country

```{sql connection=congap}
SELECT country, AVG(gdpPercap) as avgGdpPercap
FROM gapminder
GROUP BY country
```



15. Show the total population, average population, the minimum, and maximum population for Canada

```{sql connection=congap}
SELECT country, SUM(pop) as totalPop, AVG(pop) as avgPop, MIN(pop) as minPop, MAX(pop) as maxPop
FROM gapminder
WHERE country = "Canada"
```




16. From the database we also have the table countries, give the country and continent code:

```{sql connection=congap}
SELECT country, continent
FROM countries
GROUP BY country
```



17. how many countries are there per continent?

```{sql connection=congap}
SELECT continent, COUNT(country) as "# of countries"
FROM countries
GROUP BY continent
```




18. What is the population per continent? 


SELECT DISTINCT continent, pop 
FROM gapminder
INNER JOIN countries
GROUP BY continent

```{sql connection=congap}
SELECT DISTINCT continent, pop 
FROM gapminder
INNER JOIN countries
GROUP BY continent
```



## Books queries

For a second set of queries we have a database with the following database model:

<img src = "images/Books.png" />

19. Show a list of all authors whose first name starts with the ‘An’ characters.

```{sql connection=conbook}
SELECT *
FROM Authors
WHERE au_fname LIKE 'An%'
```



20. Show a list of all authors where the last 4 numbers of the id start with a 3 follow with three numbers.

```{sql connection=conbook}
SELECT *
FROM Authors
WHERE au_id LIKE '_______3___'
```



21.  list for each zip code the maximum salary and corresponding author. Only Looking at zip codes where the maximum salary is below 10.000.

```{sql connection=conbook}
SELECT DISTINCT au_id, au_salary, au_zip
FROM Authors
GROUP BY au_zip HAVING MAX(au_salary) < 10000
```



22. What is the total price for books sold from the ‘Algodata Infosystems’ publisher? Use a subquery.

```{sql connection=conbook}
SELECT SUM(b.price) as total_Price
FROM Books b INNER JOIN Publishers p
WHERE p.pub_name LIKE 'Algodata Infosystems' AND b.pub_id = p.pub_id
```



23. List all employees who are involved in publishing a book for the Binnet & Hardley publisher. (use a subquery).

```{sql connection=conbook}
SELECT e.fname, e.lname, e.emp_id
FROM Employees e INNER JOIN Publishers p
WHERE p.pub_name = 'Binnet & Hardley' AND e.pub_id = p.pub_id
```



24. What is the total number of business books that are written per gender of the author? 
List gender and the total number of books.

```{sql connection=conbook}
SELECT COUNT(b.type) as total, a.au_sex
FROM Books b JOIN Authors a JOIN BookAuthor ba
WHERE b.type LIKE "business" AND b.title_id = ba.title_id AND ba.au_id = a.au_id
GROUP BY a.au_sex
```



25. Who wrote the book with title ‘Is Anger the Enemy?’? 
List the first and the last names of the author and the author id.

```{sql connection=conbook}
SELECT a.au_fname, a.au_lname, a.au_id
FROM Books b JOIN Authors a JOIN BookAuthor ba
WHERE b.title LIKE 'Is Anger the Enemy?' AND b.title_id = ba.title_id AND ba.au_id = a.au_id
```



## Combining SQL and the tidyverse

In this section we will use the gapminder and courtries dataset. When joining these you do this on country. After making the appropriate query you can use the tidyverse library to make the plots.
Which code is used to get the below mentioned graphs?

26. compare the gdpPercap against the lifeExp for 2007 only. (use the gapminder database.)

```{sql connection=congap,output.var="DF26"}


```


```{r}
DF26
```


27. If we want to compare the countries in Oceania with each other on the growth in population over the years we could use the following graph. (use the gapminder database.)


```{sql connection=congap,output.var="DF27"}


```


```{r}
DF27
```

28. One of the reasons in growth is the life expectancy. The life expectancy is in its turn related to the gdp. To get a view for both countries we can have a look at the following graphs. Plot the lief expectancy against the gdpPercap for both countries. use the population to indicate the growth.  (use the gapminder database.)


```{sql connection=congap,output.var="DF28"}

```


```{r}
DF28
```

29. In the Books database we want to see how many books are published per state of the Authors. (use the books database.)


```{sql connection=conbook,output.var="DF29"}

```


```{r}
DF29
```

## End of test

Make sure to save everything in the right folder. 

In the submission folder you include, your notebook, databases and HTML file. 

Before you hit submit, make sure to close all other applications, including your browser and Rstudio!!

Then hit submit.